### 2014-08-20 chapter-appendix01-S1800.tex
first commit of appendix files

\chapter{S1800 photoresist technical manual}
\label{sec:A1s1800}

The next four pages are the S1800 photoresist technical manual accessed from the website \url{http://microchem.com/PDFs_Dow/S1800.pdf} on September 11, 2012.  The main features of the document that are relevant for our work are Figures 1, 4, and 7, which show the expected resist thickness, critical dose dependence on resist thickness, and an example critical dose to clear curve.

\includepdf[pages=1-4]{S1800.pdf}

\chapter{Cary 5000 Polarizer data sheet}
\label{sec:A2Cary5000Pol}

The next three pages are from the Agilent technical manual on operating the Cary 5000 in polarization mode.  I accessed the document on May 10, 2013 from the Agilent website.\footnote{\url{http://www.chem.agilent.com/en-US/products-services/instruments-systems/molecular-spectroscopy/cary-5000-uv-vis-nir/Pages/default.aspx}}

\includepdf[pages=1-4]{Cary5000pol_datasheet.pdf}