### 2015-03-25 ch1/chapter-introduction.tex
add a changelog, begin introduction text, comment out old protochapters

\chapter{Introduction}

Technology and ideas catalyze discovery.  Astronomical technology development started 405 years ago with Galileo's 2 inch telescope and \emph{Sidereus Nucius}.  Today there are sixteen optical telescopes with diameters of 6.5 meters or larger, and there are IPython Jupyter Notebooks\footnote{\url{https://jupyter.org/}}\footnote{Galileo's detailed notebooks of Jupiter contained observational data, calculations, figures, and explanatory material, and were, in part, the inspiration for the name of Jupyter Notebooks.  The \emph{py} is for python, naturally.}  This PhD dissertation  pushes the frontier of both scientific technology and observational astronomy.  There are two main threads.  First, I developed silicon diffractive optics for high resolution near infarared astronomical spectroscopy.  The devices leverage high precision silicon semiconductor lithography to provide optical quality diffraction grooves.